---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# Artesian Well Site Assessment Data

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

**Description**

This dataset provides detailed information on artesian well sites
located in Zomba, Blantyre, Mangochi, Thyolo, Lilongwe, Dowa, Balaka,
Mwanza, Neno and Chiradzulo, collected in April 2024. Each record
represents a unique well site and includes attributes related to
location (latitude and longitude), water source characteristics,
infrastructure details, site usage, environmental context, and
management structures.

Key variables

-   Geolocation: Coordinates to map and analyze spatial distribution.
-   Water source information: Whether the well is artesian, presence of
    structures, ability to test flow, and the time (in seconds) to fill
    a 20-liter container.
-   Usage patterns: Main and additional uses of the site (e.g.,
    domestic, irrigation, fish farming).
-   Nearby water abstractions: Proximity and attributes of neighboring
    abstraction points.
-   Environmental data: Seasonal availability and nearby river presence.

This dataset is complemented by images captured at each site, accessible
via permanent URLs, which provide photographic evidence of physical and
infrastructural conditions.

**Potential Use Cases**

Water Resource Planning and Management Government bodies and NGOs can
use this dataset to assess water point availability, accessibility, and
infrastructure needs.

1.  Hydrological and Environmental Studies Researchers and environmental
    scientists can analyze recharge patterns, abstraction pressures, and
    seasonal availability for groundwater sustainability assessments.
2.  Infrastructure Development and Maintenance Engineers and local
    authorities can use structure type and condition data to prioritize
    maintenance and investments in water infrastructure.
3.  Community Health and Livelihood Planning The dataset helps evaluate
    the adequacy of water sources for domestic and agricultural use,
    which is critical for public health and economic planning.
4.  Geospatial and GIS Mapping Projects The detailed coordinates and
    associated metadata support mapping efforts to visualize and analyze
    spatial patterns in water accessibility and site clustering.
5.  Monitoring and Evaluation (M&E) Organizations implementing rural
    water programs can track improvements, assess usage efficiency, and
    monitor infrastructure over time.

## Installation

You can install the development version of artesianwells from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/artesianwells")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

1.  Click Download CSV. A window opens that displays the CSV in your
    browser.
2.  Right-click anywhere inside the window and select "Save Page As...".
3.  Save the file in a folder of your choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/openwashdata/artesianwells/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to detailed information on artesian well
sites

```{r}
library(artesianwells)
```

### artesianwells

The dataset `artesianwells` contains `r nrow(artesianwells)`
observations and `r ncol(artesianwells)` variables

```{r}
artesianwells |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "artesianwells.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Example

```{r}
library(artesianwells)
library(dplyr)
library(ggplot2)
library(maps)
# Filter usable water point locations
water_map_data <- artesianwells %>%
  filter(!is.na(latitude) & !is.na(longitude))  # Ensure geo-points are valid

# Get Malawi map outline
malawi_map <- map_data("world", region = "Malawi")

# Plot the water point locations
ggplot() +
  geom_polygon(data = malawi_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = "gray70") +
  geom_point(data = water_map_data,
             aes(x = longitude, y = latitude, color = use_of_site),
             size = 2, alpha = 0.7) +
  coord_fixed(1.3) +
  labs(
    title = "Artesian Wells in Malawi",
    x = "Longitude", y = "Latitude", color = "Use of Water Site"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

```

## License

Data are available as
[CC-BY](https://github.com/openwashdata/%7B%7B%7Bpackagename%7D%7D%7D/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("artesianwells")
```
