---
title: "Artesian Wells Image Gallery"
author:
  - name: Lars Schöbitz
    email: lschoebitz@ethz.ch
    orcid: 0000-0003-2196-5015
    affiliation:
      - name: Global Health Engineering, ETH Zurich
        url: https://ghe.ethz.ch
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
execute:
  message: false
  warning: false
---

```{r setup}
#| include: false
library(artesianwells)
library(tidyverse)
library(gt)
library(gtExtras)
```

## Introduction

This vignette provides a visual showcase of the artesian wells documented in the dataset. The images are organized by district and displayed alongside key information about each well site.

```{r data-prep}
# Prepare data with image URLs
wells_with_images <- artesianwells %>%
  filter(!is.na(well_images)) %>%
  mutate(
    # Extract first image URL if multiple are separated by semicolons
    first_well_image = str_extract(well_images, "^[^;]+"),
    first_area_photo = if_else(!is.na(area_photos), 
                               str_extract(area_photos, "^[^;]+"), 
                               NA_character_)
  ) %>%
  select(
    district,
    traditional_authority,
    latitude,
    longitude,
    has_structure,
    well_structure_type,
    seconds_to_fill_20L,
    use_of_site,
    river_nearby,
    has_operator,
    mgmt_type,
    first_well_image,
    first_area_photo
  )

# Count wells by district
district_summary <- wells_with_images %>%
  group_by(district) %>%
  summarise(
    well_count = n(),
    avg_fill_time = round(mean(seconds_to_fill_20L, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(desc(well_count))
```

## Overview

The dataset contains images for **`r nrow(wells_with_images)`** artesian well sites across **`r n_distinct(wells_with_images$district)`** districts in Malawi.

```{r overview-table}
#| label: tbl-district-summary
#| tbl-cap: "Summary of artesian wells by district"

district_summary %>%
  gt() %>%
  cols_label(
    district = "District",
    well_count = "Number of Wells",
    avg_fill_time = "Avg. Fill Time (seconds)"
  ) %>%
  tab_header(
    title = "Artesian Wells Distribution",
    subtitle = "Wells with documented images"
  ) %>%
  fmt_number(
    columns = avg_fill_time,
    decimals = 1
  ) %>%
  tab_style(
    style = cell_fill(color = "lightblue"),
    locations = cells_body(columns = well_count, rows = well_count > 5)
  )
```

## Wells by District

```{r district-galleries}
#| output: asis
#| echo: false

# Function to create a district section
create_district_section <- function(district_name, data) {
  cat("\n\n### ", district_name, "\n\n")
  
  district_data <- data %>%
    filter(district == district_name) %>%
    arrange(traditional_authority)
  
  # Create the table with images
  table_html <- district_data %>%
    mutate(
      location = paste0(round(latitude, 4), ", ", round(longitude, 4)),
      fill_rate = case_when(
        is.na(seconds_to_fill_20L) ~ "Not tested",
        seconds_to_fill_20L < 30 ~ "Fast (< 30s)",
        seconds_to_fill_20L < 60 ~ "Moderate (30-60s)",
        TRUE ~ "Slow (> 60s)"
      )
    ) %>%
    select(
      first_well_image,
      traditional_authority,
      location,
      has_structure,
      fill_rate,
      use_of_site,
      has_operator
    ) %>%
    gt() %>%
    cols_label(
      first_well_image = "Well Image",
      traditional_authority = "Traditional Authority",
      location = "Coordinates",
      has_structure = "Structure",
      fill_rate = "Fill Rate",
      use_of_site = "Primary Use",
      has_operator = "Operator"
    ) %>%
    tab_header(
      title = paste0("Artesian Wells in ", district_name),
      subtitle = paste0(nrow(district_data), " well sites documented")
    ) %>%
    fmt_image(
      columns = first_well_image,
      height = 150,
      width = 150
    ) %>%
    cols_width(
      first_well_image ~ px(160),
      traditional_authority ~ px(150),
      location ~ px(120),
      has_structure ~ px(80),
      fill_rate ~ px(100),
      use_of_site ~ px(100),
      has_operator ~ px(80)
    ) %>%
    tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels()
    ) %>%
    tab_style(
      style = cell_fill(color = "lightgreen"),
      locations = cells_body(
        columns = fill_rate,
        rows = fill_rate == "Fast (< 30s)"
      )
    ) %>%
    tab_style(
      style = cell_fill(color = "lightyellow"),
      locations = cells_body(
        columns = fill_rate,
        rows = fill_rate == "Moderate (30-60s)"
      )
    ) %>%
    tab_style(
      style = cell_fill(color = "lightcoral"),
      locations = cells_body(
        columns = fill_rate,
        rows = fill_rate == "Slow (> 60s)"
      )
    ) %>%
    as_raw_html()
  
  cat(table_html)
}

# Create sections for each district
for (dist in district_summary$district) {
  create_district_section(dist, wells_with_images)
}
```

## Image Grid View

For a quick visual overview, here's a grid displaying all well images organized in rows of 4:

```{css}
#| echo: false
.image-grid {
  display: grid;
  grid-template-columns: repeat(4, 1fr);
  gap: 10px;
  margin: 20px 0;
}

.image-item {
  text-align: center;
}

.image-item img {
  width: 100%;
  height: 150px;
  object-fit: cover;
  border-radius: 5px;
  box-shadow: 0 2px 4px rgba(0,0,0,0.1);
}

.image-caption {
  font-size: 0.8em;
  margin-top: 5px;
  color: #666;
}

@media (max-width: 768px) {
  .image-grid {
    grid-template-columns: repeat(2, 1fr);
  }
}

@media (max-width: 480px) {
  .image-grid {
    grid-template-columns: 1fr;
  }
}
```

::: {.image-grid}
```{r}
#| output: asis
#| echo: false

# Create image grid
wells_with_images %>%
  mutate(
    caption = paste0(district, " - ", traditional_authority)
  ) %>%
  select(first_well_image, caption) %>%
  slice_head(n = 40) %>%  # Show first 40 images
  pwalk(function(first_well_image, caption) {
    cat(paste0(
      '<div class="image-item">',
      '<img src="', first_well_image, '" alt="', caption, '">',
      '<div class="image-caption">', caption, '</div>',
      '</div>\n'
    ))
  })
```
:::

## Usage Patterns

```{r usage-analysis}
#| label: fig-usage-distribution
#| fig-cap: "Distribution of artesian wells by primary use"

wells_with_images %>%
  count(use_of_site) %>%
  mutate(use_of_site = fct_reorder(use_of_site, n)) %>%
  ggplot(aes(x = n, y = use_of_site, fill = use_of_site)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Number of Wells",
    y = "Primary Use",
    title = "Primary Usage of Artesian Wells"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Flow Rate Analysis

```{r flow-analysis}
#| label: tbl-flow-stats
#| tbl-cap: "Flow rate statistics by district"

wells_with_images %>%
  filter(!is.na(seconds_to_fill_20L)) %>%
  group_by(district) %>%
  summarise(
    n_tested = n(),
    min_time = min(seconds_to_fill_20L, na.rm = TRUE),
    avg_time = mean(seconds_to_fill_20L, na.rm = TRUE),
    max_time = max(seconds_to_fill_20L, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gt() %>%
  cols_label(
    district = "District",
    n_tested = "Wells Tested",
    min_time = "Min (s)",
    avg_time = "Average (s)",
    max_time = "Max (s)"
  ) %>%
  fmt_number(
    columns = c(min_time, avg_time, max_time),
    decimals = 1
  ) %>%
  tab_header(
    title = "Flow Rate Statistics",
    subtitle = "Time to fill 20-liter container"
  ) %>%
  data_color(
    columns = avg_time,
    colors = scales::col_numeric(
      palette = c("green", "yellow", "red"),
      domain = NULL
    )
  )
```

## Data Notes

- Images are hosted on the mwater API and may require internet connection to display
- Some wells have multiple images; only the first image is displayed in this gallery
- Flow rate data is not available for all wells
- The dataset was collected in April 2024